Eco-evolutionary dynamics of finite populations from first principles

### 2023-03-24 ind_files/05_unification_discussion.tex
Minor rephrasing

More generally, the well-known intimate relation between It\^o SDEs and Fokker-Planck equations (see section \ref{sec_math_background}) is extremely useful and has been extensively exploited in the literature~\citep{van_kampen_stochastic_1981,oksendal_stochastic_1998, gardiner_stochastic_2009}, because some problems are much more easily attacked using tools from probability theory that operate on It\^o processes, whereas others are more amenable to study using tools from PDE theory that operate on the Fokker-Planck equations. In the infinite-dimensional case, working with the field theory through functional Fokker-Planck equations such as \eqref{disc_functional_field_eqns} allows us to use tools with an essentially `differential equation' flavor such as spectral methods (\cite{rogers_demographic_2012, rogers_modes_2015}; See also Appendix \ref{App_examples} for a general model-independent pedagogical example), fast-mode (`adiabatic') elimination/slow manifold approximation~\citep{constable_stochastic_2013,parsons_dimension_2017}, and WKB style expansions to detect extreme events~(\cite{rogers_demographic_2012,assaf_wkb_2017}; though it bears noting that WKB theory is closely related to large deviation principles that occur in the study of Markov processes and often come in a distinctly `probabilistic' flavor, see~\cite{dembo_large_1998}). On the other hand, many biologically interesting questions are much easier to approach from the branching processes/SPDE side using tools that have an essentially `probabilistic' flavor such as duality~\citep{dawson_stochastic_1975,greenman_duality_2020}, infinitesimal generators/semi-group theory~\citep{ethier_markov_1986,finkelshtein_semigroup_2012}, and the martingale problem~\citep{ethier_markov_1986,dawson_stochastic_2000, champagnat_unifying_2006}. Rigorously establishing and systematically exploiting relations between functional Fokker-Planck equations of the form \eqref{disc_functional_field_eqns} and the sort of general SPDEs studied by~\citet{week_white_2021} for measure-valued branching processes could thus prove very fruitful for developing more integrative eco-evolutionary theory since it allows us to seamlessly transition between complementary views of the same object to suit the problem at hand\footnote{The physics literature has long recognized this. Physicists regularly exploit the connections between stochastic/quantum field equations and SPDEs, often studying SPDEs in completely `classical' systems using tools and language from quantum field theory~\citep{hochberg_effective_1999,baez_quantum_2018}.}.

>>>>

More generally, the well-known intimate relation between It\^o SDEs and Fokker-Planck equations (see section \ref{sec_math_background}) is extremely useful and has been extensively exploited in the literature~\citep{van_kampen_stochastic_1981,oksendal_stochastic_1998, gardiner_stochastic_2009}, because some problems are much more easily attacked using tools from probability theory that operate on It\^o processes, whereas others are more amenable to study using tools from PDE theory that operate on the Fokker-Planck equations. In the infinite-dimensional case, working with the field theory through functional Fokker-Planck equations such as \eqref{disc_functional_field_eqns} allows us to use tools with an essentially `differential equation' flavor such as spectral methods (\cite{rogers_demographic_2012, rogers_modes_2015}; See also Appendix \ref{App_examples} for a general model-independent pedagogical example), fast-mode (`adiabatic') elimination/slow manifold approximation~\citep{constable_stochastic_2013,parsons_dimension_2017}, and WKB style expansions to detect extreme events~(\cite{rogers_demographic_2012,assaf_wkb_2017}; though it bears noting that WKB theory is closely related to large deviation principles that occur in the study of Markov processes and often come in a distinctly `probabilistic' flavor, see~\cite{dembo_large_1998}). On the other hand, many biologically interesting questions are much easier to approach from the branching processes/SPDE side using tools that have an essentially `probabilistic' flavor such as duality~\citep{dawson_stochastic_1975,greenman_duality_2020}, infinitesimal generators/semi-group theory~\citep{ethier_markov_1986,finkelshtein_semigroup_2012}, and the martingale problem~\citep{ethier_markov_1986,dawson_stochastic_2000, champagnat_unifying_2006,champagnat_individual_2008}. Rigorously establishing and systematically exploiting relations between functional Fokker-Planck equations of the form \eqref{disc_functional_field_eqns} and the sort of general SPDEs studied by~\citet{week_white_2021} for measure-valued branching processes could thus prove very fruitful for developing more integrative eco-evolutionary theory since it allows us to seamlessly transition between complementary views of the same object to suit the problem at hand\footnote{The physics literature has long recognized this. Physicists regularly exploit the connections between stochastic/quantum field equations and SPDEs, often studying SPDEs in completely `classical' systems using tools and language from quantum field theory~\citep{hochberg_effective_1999,baez_quantum_2018}.}.
